---
output:
  pdf_document: default
  html_document: default
---

# Power Analysis 

```{r declare_constants_and_libraries, include = FALSE, warning = FALSE}

# install.packages('lmtest')
library(data.table)
library(dplyr)
library(ggplot2)
library(magrittr)
library(lmtest)
library(sandwich)

# Significance threshold
ALPHA = 0.05
 
# Profile_Viewer_Attract
# Figures based on "The Democracy of Dating"
ATTRACT_CONTROL = 89
ATTRACT_LIBERAL = 83
ATTRACT_CONSERVATIVE = 72

# Returns percent difference (as a decimal) compared to the old
get_pct_diff <- function(old, new) {
  return((new-old)/old)
}

# Treatment multipliers based on percentage less likely to swipe
TRT_LIB = get_pct_diff(ATTRACT_CONTROL, ATTRACT_LIBERAL)
TRT_CON = get_pct_diff(ATTRACT_CONTROL, ATTRACT_CONSERVATIVE)
# Base treatment modifier
TRT_LIB_SD_PCT = 0.01
TRT_CON_SD_PCT = 0.01
```

The research question investigated in this study is: "Does declaring a political leaning in a dating application profile affect the profile's match rate?". To answer this question, this study will attempt to create dating profiles of different genders in different dating application platforms with biography descriptions containing one of three treatment variants: No political leaning (Control), Liberal leaning (Treatment 1), or Conservative leaning (Treatment 2).</br> 

The profiles will be exposed with no treatment added to the user network of the dating application platform. All profiles will attempt to match with the first designated number of profiles shown to them, then number of matches within the first testing period will be measured. After a break period, the treatment will be administered to the two treatment groups and the profiles will attempt to be match with the next fixed number of profiles shown to them. The number of matches will be recorded and measured for the treatment effect. </br>  

To measure the treatment effect, three scenarios were simulated in this Power Analysis report:
1. The nominal number of matches a profile receives after the treatment will be measured and compared between treatment groups for any treatment effect.
2. The number of post-treatment matches as a percentage of the pre-treatment matches (`post-treatment matches` / `pre-treatment matches`) will be measured and compared between treatment groups.
3. 

A data table was created to simulate the population and simulate the treatment effects on different profiles where each row represents an experimental dating profile created and exposed to the dating profile network with or without a treatment. The following considerations were made:

* 75% of the profiles generated were male and 25% were female according to studies
* The nominal number of matches per profile between three dating platforms (Tinder, Hinge, Plenty of Fish) were scaled by market share as studied by [Businessofapps.com](https://www.businessofapps.com/data/dating-app-market/)
* The estimated population treatment effects were derived from findings from [The Democracy of Dating: A Survey Experiment on American Dating Preferences](https://scholarsarchive.byu.edu/cgi/viewcontent.cgi?article=1236&context=sigma) by Matthew Easton where the overall attractiveness of Liberal profiles was 6.74% less than that of control and the overall attractiveness of Conservative profiles was 19.1% less than that of control. These percentage drops were incorporated in the treatment effect.
* The treatment effects and figures were randomized using normal distributions

```{r create_table, warning = FALSE}
# Use 30k rows as the population
d <- data.table(id = 1:30000)

# assign participants a gender and app platform
d[, gender := sample(c('male', 'female'), size = .N, replace = TRUE)]
d[, platform := sample(c('tinder', 'hinge', 'pof'), size = .N, replace = TRUE)]

# simulates variation due to different profile pictures
d[, quality_multiplier := pmax(rnorm(.N,1.5,0.25),0.01)]

GENDER_MALE_FEMALE_RATIO <- 3
GENDER_MALE_SD_PCT <- 0 # 0.1
GENDER_FEMALE_SD_PCT <- 0 # 0.1

# https://www.businessofapps.com/data/dating-app-market/
PLATFORM_TINDER_MARKETSHARE <- 32
PLATFORM_HINGE_MARKETSHARE <- 15
PLATFORM_POF_MARKETSHARE <- 14

PLATFORM_TINDER_SD_PCT <- 0 # 0.1
PLATFORM_HINGE_SD_PCT <- 0 # 0.1
PLATFORM_POF_SD_PCT <- 0 # 0.1

# studies show dating apps are ~75% male
d[gender == 'male',     gender_multiplier := pmax(rnorm(.N,1,1*GENDER_MALE_SD_PCT),0.01)]
d[gender == 'female',     gender_multiplier := pmax(rnorm(.N,GENDER_MALE_FEMALE_RATIO,GENDER_MALE_FEMALE_RATIO*GENDER_FEMALE_SD_PCT),0.01)]

d[platform == 'tinder',     audience_multiplier := pmax(rnorm(.N,PLATFORM_TINDER_MARKETSHARE,PLATFORM_TINDER_MARKETSHARE*PLATFORM_TINDER_SD_PCT),0.01)]
d[platform == 'hinge',     audience_multiplier := pmax(rnorm(.N,PLATFORM_HINGE_MARKETSHARE,PLATFORM_HINGE_MARKETSHARE*PLATFORM_HINGE_SD_PCT),0.01)]
d[platform == 'pof',     audience_multiplier := pmax(rnorm(.N,PLATFORM_POF_MARKETSHARE,PLATFORM_POF_MARKETSHARE*PLATFORM_POF_SD_PCT),0.01)]

# simulates pool of number of matches before treatment
d[,base_matches := round(quality_multiplier * gender_multiplier * audience_multiplier)]

# different treatments to be assigned
d[, treatment := sample(c('anone', 'liberal', 'conservative'), size = .N, replace = TRUE)]

# Treatment-induced decay figures based on "The Democracy of Dating"
d[treatment == 'anone',     treatment_multiplier := 0]
# propose liberal treatment causes 15% fewer matches on average
d[treatment == 'liberal',     treatment_multiplier := pmax(rnorm(.N,TRT_LIB,abs(TRT_LIB)*TRT_LIB_SD_PCT),-0.99)]
# propose conservative treatment causes 30% fewer matches on average
d[treatment == 'conservative',     treatment_multiplier := pmax(rnorm(.N,TRT_CON,abs(TRT_CON)*TRT_CON_SD_PCT),-0.99)]

# simulate natural decay in number of matches relative to when account is first created
d[, time_decay_multiplier := pmax(rnorm(.N,0.75,0.05),0.01)]

# simulates pool of number of matches after treatment (or control)
d[,num_matches := round(base_matches * time_decay_multiplier * (1 + treatment_multiplier))]
# measure the number of matches after treatment relative to number of matches during the pre-treatment control
d[,rel_match_rate := num_matches/base_matches]

```


``` {r random_inference_functions, warning = FALSE}
# Samples from the population table and secures equal number of
# profiles for control and both treatments (must be in multiples of 3)
get_sample <- function(d, s_size = 18) {
  return(d[, .SD[sample(x = .N, size = s_size/3,replace=TRUE)], by = treatment])
  # return(d[, .SD[sample(x = .N, size = s_size/18,replace=TRUE)], by = .(treatment,gender,platform)])
}

# Returns the T-Test object of the difference between means
# Compares Liberal-Control or Conservative-Control
experiment_t_test <- function(d_sample,  column, is_liberal=TRUE) {
  
  # Separate into control and treatment
  v_treatment <- NA
  v_control <- NA
  
  # Selects Liberla or Conservative comparison
  
  # if ( TRUE == is_liberal ) {
  #   d_sample = d_sample[treatment != 'conservative',]
  # } else {
  #   d_sample = d_sample[treatment != 'liberal',]
  # }
  # 
  # if (column == "num_matches") {
  #   return(t.test(num_matches ~ treatment,data=d_sample))
  # } else {
  #   return(t.test(rel_match_rate ~ treatment,data=d_sample))
  # }
  
  treat = NA
  if (TRUE == is_liberal) {
    treat = 'liberal'
  } else {
    treat = 'conservative'
  }

  if (column == "num_matches") { # Scenario 1
    v_treatment <- d_sample[treatment == treat, "num_matches"]
    v_control <- d_sample[treatment == 'anone', "num_matches"]
  } else if (column == "rel_match_rate") { # Scenario 2
    v_treatment <- d_sample[treatment == treat, "rel_match_rate"]
    v_control <- d_sample[treatment == 'anone', "rel_match_rate"]
  } else if (column == "matched") { # Scenario 3
    # Add column name here
    v_treatment <- d_sample[treatment == treat, "matched"]
    v_control <- d_sample[treatment == 'anone', "matched"]
  }

  # Run T-test
  t_test_obj = t.test(v_control, v_treatment, conf.level = 0.95, alternative = "two.sided")

  return(t_test_obj)
}

# Returns only the p-value from the experiment() function
experiment_t_test_p_value_only <- function(d_sample, column, is_liberal=TRUE) {
  return(experiment_t_test(d_sample, column, is_liberal)$"p.value")
}

experiment_multi_treatment_t_test <- function(d, s_size, column) {
  # Get sample
  d_sample <- get_sample(d, s_size)
  l_treat <- experiment_t_test(d_sample, column=column, is_liberal = TRUE)
  c_treat <- experiment_t_test(d_sample, column=column, is_liberal = FALSE)
  return(c(l_treat, c_treat))
}

experiment_multi_treatment_t_test_p_value_only <- function(d, s_size, column) {
  # Get sample
  d_sample <- get_sample(d, s_size)

  l_treat <- experiment_t_test_p_value_only(d_sample, column=column, is_liberal = TRUE)
  c_treat <- experiment_t_test_p_value_only(d_sample, column=column, is_liberal = FALSE)
  return(c(l_treat, c_treat))
}

experiment_lm <- function(d_sample,  column) {

  if (column == "num_matches") { # Scenario 1
    return(lm(log(num_matches+1) ~ factor(treatment) + factor(gender) + factor(platform),data=d_sample))
    # return(lm(log(num_matches+1) ~ factor(treatment),data=d_sample))
  } else if (column == "rel_match_rate") { # Scenario 2
    return(lm(log(rel_match_rate+0.001) ~ factor(treatment) + factor(gender) + factor(platform),data=d_sample))
    # return(lm(log(rel_match_rate+0.001) ~ factor(treatment),data=d_sample))
  } else if (column == "matched") { # Scenario 3
    return(lm(matched ~ factor(treatment)  + factor(gender),data=d_sample))
  }
}

experiment_lm_p_value_only <- function(d, s_size,  column) {
  d_sample <- get_sample(d, s_size)
  e_result <- experiment_lm(d_sample,column)
  
  # c_test <- coeftest(e_result)
  # use robust errors
  c_test <- coeftest(e_result,vcov.=vcovHC(e_result))
  return(c(c_test['factor(treatment)liberal',4],c_test['factor(treatment)conservative',4]))
}

# Gets statistical power (percentage of tests that reject the null hypothesis)
# Returns two (2) values: Liberal-Control power & Conservative-Control power
get_power <- function(d,n_experiments, s_size=9, column, method='lm') {
  t_test_p_values_lib <- c()
  t_test_p_values_con <- c()

  for (i in 1:n_experiments) {
    if (method == 't.test') {
      treat_p_values <- experiment_multi_treatment_t_test_p_value_only(d,s_size,column)
      t_test_p_values_lib[i] <- treat_p_values[1]
      t_test_p_values_con[i] <- treat_p_values[2]
    } else {
      treat_p_values <- experiment_lm_p_value_only(d,s_size,column)
      t_test_p_values_lib[i] <- treat_p_values[1]
      t_test_p_values_con[i] <- treat_p_values[2]
    }
  }
  
  t_test_rejects_lib <- t_test_p_values_lib[t_test_p_values_lib <= ALPHA]
  pct_reject_lib = length(t_test_rejects_lib)/n_experiments*100
  
  t_test_rejects_con <- t_test_p_values_con[t_test_p_values_con <= ALPHA]
  pct_reject_con = length(t_test_rejects_con)/n_experiments*100
  
  return(c(pct_reject_lib, pct_reject_con))
}

# Scenario 1: Use raw match rate
#mod_1 <- lm(num_matches ~ factor(treatment) + factor(platform) + factor(gender),data=d)

# Scenario 2: Use match decay
#mod_2 <- lm(rel_match_rate ~ factor(treatment) + factor(platform) + factor(gender),data=d)

# Scenario 3: Use


#coeftest(mod_1)

profile_sample_sizes <-  seq(18, 180, by = 18)
swipe_sample_sizes <- seq(18, 180, by = 18)

test_scenario <- function(d, n_experiments,sample_sizes,column='num_matches', method='lm') {
  powers_lib = c()
  powers_con = c()
  count = 1
  for (s in sample_sizes) {
    powers = get_power(d, n_experiments = n_experiments, s_size = s, column = column, method = method)
    powers_lib[count] = powers[1]
    powers_con[count] = powers[2]
    count = count + 1
  }

  return(list(powers_lib, powers_con))
}

scenario_1 = test_scenario(d,1000,profile_sample_sizes, column='num_matches',method='lm')
scenario_2 = test_scenario(d,1000,profile_sample_sizes, column='rel_match_rate',method='lm')
scenario_3 = test_scenario(dswipe,1000, swipe_sample_sizes, column='matched',method='lm')

```

```{r plot_results, warning = FALSE}

dt_1 <- data.table("s_size" = profile_sample_sizes, "s1_lib" = unlist(scenario_1[1]), "s1_con" = unlist(scenario_1[2]), "s2_lib" = unlist(scenario_2[1]), "s2_con" = unlist(scenario_2[2]))


colors_1 <- c(
  "Nominal Match Rate - Liberal" = "blue",
  "Nominal Match Rate - Conservative" = "red",
  "Relative Match Rate - Liberal" = "darkblue",
  "Relative Match Rate - Conservative" = "darkred"
  )


ggplot(dt_1, aes(x=profile_sample_sizes)) + 
  geom_line(aes(y = s1_lib, color = "Nominal Match Rate - Liberal"), linetype="dashed") + 
  geom_line(aes(y = s1_con, color= "Nominal Match Rate - Conservative"), linetype="dashed") +
  geom_line(aes(y = s2_lib, color = "Relative Match Rate - Liberal"), linetype="twodash") + 
  geom_line(aes(y = s2_con, color= "Relative Match Rate - Conservative"), linetype="twodash") +
  scale_colour_manual(values = colors_1) +
  labs(x="Number of Profiles", y="(%)", color = "Legend", title = 'LM No covar')

```

```{r}
dswipe <- data.table(id = 1:300000)

PCT_FEMALE_SWIPE_MATCH <- 0.19
PCT_MALE_SWIPE_MATCH <-0.055

ATTRACT_LIBERAL <- 50
ATTRACT_CONSERVATIVE <- 30

dswipe[, gender := sample(c('male', 'female'), size = .N, replace = TRUE)]
dswipe[, treatment := sample(c('anone', 'liberal', 'conservative'), size = .N, replace = TRUE)]

dswipe[treatment == 'anone' &        gender == 'male',   matched := rbinom(.N,1,PCT_MALE_SWIPE_MATCH)]
dswipe[treatment == 'anone' &        gender == 'female', matched := rbinom(.N,1,PCT_FEMALE_SWIPE_MATCH)]

dswipe[treatment == 'liberal' &     gender == 'male',   matched := rbinom(.N,1,PCT_MALE_SWIPE_MATCH*ATTRACT_LIBERAL/ATTRACT_CONTROL)]
dswipe[treatment == 'liberal' &     gender == 'female', matched := rbinom(.N,1,PCT_FEMALE_SWIPE_MATCH*ATTRACT_LIBERAL/ATTRACT_CONTROL)]

dswipe[treatment == 'conservative' & gender == 'male',   matched := rbinom(.N,1,PCT_MALE_SWIPE_MATCH*ATTRACT_CONSERVATIVE/ATTRACT_CONTROL)]
dswipe[treatment == 'conservative' & gender == 'female', matched := rbinom(.N,1,PCT_FEMALE_SWIPE_MATCH*ATTRACT_CONSERVATIVE/ATTRACT_CONTROL)]


```

```{r}

dt_2 <- data.table("s_size" = swipe_sample_sizes,"s3_lib" = unlist(scenario_3[1]), "s3_con" = unlist(scenario_3[2]))


colors_2 <- c(
  "S3 Liberal" = "lightblue",
  "S3 Conservative" = "pink"
  )


ggplot(dt_2, aes(x=swipe_sample_sizes)) + 
  geom_line(aes(y = s3_lib, color = "S3 Liberal"), linetype="twodash") + 
  geom_line(aes(y = s3_con, color = "S3 Conservative"), linetype="twodash") +
  scale_colour_manual(values = colors_2) +
  labs(x="Number of Swipes", y="(%)", color = "Legend", title = 'LM No covar')

```

Based on the analysis, it seems that Scenarios 1 and 2 (nominal post-treatment match rate and post-treatment match rate relative to the number of pre-treatment matches respectively) had similar power where the estimated treatment effect of conservative profiles seemed to have much more statistical power than that of liberal profiles.


